---
title: "GI0"
output: 
  rmarkdown::html_vignette:
    fig_width: 6
    fig_height: 4
    dpi: 150
vignette: >
  %\VignetteIndexEntry{GI0}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(GI0)

```
<h1> 1. Graphical Examples</h1>

Graphical examples to explore Generalized Inverse Gaussian (GI0) distribution using density plots, cumulative distribution function plots, and quantile function plots.

<h2>1.1. Density function</h2>

The density function of the GI0 distribution is used to compute the density values for different parameter settings of alpha, gamma, and the number of looks.

```{r, warning=FALSE}
# Example of density plots
# a=-1, g=1, L=1 semilogarithmic scale: it is not an exponential density
plot(function(x) dGI0(x, p_alpha=-1, p_gamma=1, p_Looks=1), from=0, to=10, 
     log="y", type="l", xlab="x", ylab="density",
     main="Density (a=-1, g=1, L=1)")
```
```{r, warning=FALSE}
# Checking the effect of changing the number of looks
x_vals <- seq(0, 20, length.out = 200)
density_L1 <- sapply(x_vals, function(x) dGI0(x, p_alpha = -1, p_gamma = 1, p_Looks = 1))
density_L2 <- sapply(x_vals, function(x) dGI0(x, p_alpha = -1, p_gamma = 1, p_Looks = 2))
density_L10 <- sapply(x_vals, function(x) dGI0(x, p_alpha = -1, p_gamma = 1, p_Looks = 10))

plot(x_vals, density_L1, type = "l", col = "darkred", xlab = "x", ylab = "densities",
     main = "Densities changing Number of Looks", lwd=2)
lines(x_vals, density_L2, col = "coral3", lwd=2)
lines(x_vals, density_L10, col = "aquamarine4", lwd=2)
legend("topright", legend = c("L=1", "L=2", "L=10"), col = c("darkred", "coral3", "aquamarine4"), lty = 1)
```

```{r, warning=FALSE}

# Checking effect of alpha

x <- seq(0, 10, length.out = 100) 
plot(x, log="y", type="n",xlim=c(0,1), xlab="x", ylab="densities", main="Densities changing Alpha")
lines(x, dGI0(x, p_alpha=-1, p_gamma=1, p_Looks=1), col="skyblue4", lty=1,lwd=2)
lines(x, dGI0(x, p_alpha=-3, p_gamma=1, p_Looks=1), col="coral3", lty=1,lwd=2)
lines(x, dGI0(x, p_alpha=-10, p_gamma=1, p_Looks=1), col="aquamarine2", lty=1,lwd=2)
legend("topright", legend=c("a=-1", "a=-3", "a=-10"), col=c("skyblue4", "coral3", "aquamarine2"), lty=1)

```
```{r, warning=FALSE}
# Checking the effect of changing gamma
plot(x, dGI0(x, p_alpha=-3, p_gamma=1, p_Looks=1),  
     log="y", type="l", xlab="x", ylab="densities",
     main="Densities changing Gamma")
lines(x, dGI0(x, p_alpha=-3, p_gamma=3, p_Looks=1), col="coral2")
lines(x, dGI0(x, p_alpha=-3, p_gamma=10, p_Looks=1), col="aquamarine3")
legend("topright", legend=c("g=1", "g=3", "g=10"), col=c("skyblue3",
The cumulative distribution function (CDF) is shown in the following plots with the effects of changing parameters: log.p, p_Looks, p_alpha, p_g "coral2", "aquamarine3"), lty=1)
```
<h2>1.2. Cumulative distribution function</h2>
amma.

```{r, warning=FALSE}
# Checking log.p
x <- seq(0, 10, length.out = 100) 
plot(x, log="y", type="n", xlim=c(0,2), ylab="Cumulative Distribution",
     main="Cumulative Distribution Function (a=-3, g=1, L=1)")

lines(x, pGI0(x, p_alpha=-3, p_gamma=1, p_Looks=1), col="coral3", lty=1,lwd=2)
lines(x, pGI0(x, p_alpha=-3, p_gamma=1, p_Looks=1, log.p=TRUE), col="aquamarine2", lty=1,lwd=2)
legend("topright", legend=c("a=-3", "a=-10"), col=c("coral3", "aquamarine2"), lty=1)

```
```{r, warning=FALSE}
# Checking number of looks
x_vals <- seq(0, 20, length.out = 200)
cdf_L1 <- sapply(x_vals, function(x) pGI0(x, p_alpha = -1, p_gamma = 1, p_Looks = 1))
cdf_L2 <- sapply(x_vals, function(x) pGI0(x, p_alpha = -1, p_gamma = 1, p_Looks = 2))
cdf_L10 <- sapply(x_vals, function(x) pGI0(x, p_alpha = -1, p_gamma = 1, p_Looks = 10))

plot(x_vals, cdf_L1, type = "l", col = "darkred", xlab = "x", ylab = "Cumulative distribution",
     main = "CDF changing Number of Looks", lwd=2)
lines(x_vals, cdf_L2, col = "coral3", lwd=2)
lines(x_vals, cdf_L10, col = "aquamarine4", lwd=2)
legend("topright", legend = c("L=1", "L=2", "L=10"), col = c("darkred", "coral3", "aquamarine4"), lty = 1)


```
```{r, warning=FALSE}
# Checking alpha
x <- seq(0, 10, length.out = 100) 
plot(x, log="y", type="n", xlim=c(0,2), ylab="Cumulative Distribution",
     main="Cumulative Distribution Function (a=-3, g=1, L=1)")
lines(x, pGI0(x, p_alpha=-1, p_gamma=1, p_Looks=1), col="skyblue4", lty=1,lwd=2)
lines(x, pGI0(x, p_alpha=-3, p_gamma=1, p_Looks=1), col="coral3", lty=1,lwd=2)
lines(x, pGI0(x, p_alpha=-10, p_gamma=1, p_Looks=1), col="aquamarine2", lty=1,lwd=2)
legend("topright", legend=c("a=-1", "a=-3", "a=-10"), col=c("skyblue4", "coral3", "aquamarine2"), lty=1)
```
```{r, warning=FALSE}
# Checking gamma
plot(x, pGI0(x, p_alpha=-3, p_gamma=1, p_Looks=1),  
     log="y", type="l", xlab="x", ylab="Cumulative distribution",
     main="CDF changing Gamma")
lines(x, pGI0(x, p_alpha=-3, p_gamma=3, p_Looks=1), col="coral2")
lines(x, pGI0(x, p_alpha=-3, p_gamma=10, p_Looks=1), col="aquamarine3")
legend("topright", legend=c("g=1", "g=3", "g=10"), col=c("skyblue3", "coral2", "aquamarine3"), lty=1)
```
<h2>1.3. Quantile function</h2>

The quantile function is visualized in the following plots with the following parameters  changed. 

```{r, warning=FALSE}
# Checking lower.tail
x <- seq(0, 1, length.out = 100) 
plot(x, type="n", xlim=c(0,2), ylab="Quantiles",
     main="Quantile chaking lower tail")
lines(x, qGI0(x, p_alpha=-3, p_gamma=1, p_Looks=1), col="coral3", lty=1,lwd=2)
lines(x, qGI0(x, p_alpha=-3, p_gamma=1, p_Looks=1, lower.tail=FALSE), col="aquamarine2", lty=1,lwd=2)
legend("topright", legend=c("a=-3", "a=-10"), col=c("coral3", "aquamarine2"), lty=1)

```

```{r, warning=FALSE}
# Checking alpha
x <- seq(0, 1, length.out = 100) 
plot(x, log="y", type="n", xlim=c(0,2), ylab="Quantiles",
     main="Quantile changing alpha")
lines(x, qGI0(x, p_alpha=-1, p_gamma=1, p_Looks=1), col="skyblue4", lty=1,lwd=2)
lines(x, qGI0(x, p_alpha=-3, p_gamma=1, p_Looks=1), col="coral3", lty=1,lwd=2)
lines(x, qGI0(x, p_alpha=-10, p_gamma=1, p_Looks=1), col="aquamarine2", lty=1,lwd=2)
legend("topright", legend=c("a=-1", "a=-3", "a=-10"), col=c("skyblue4", "coral3", "aquamarine2"), lty=1)

```
<h2>2. Histograms and Simulated Data</h2>

rGI0 function is used to simulate data from the GI0 distribution and create histograms to visualize the data distribution with the underlying density. Freedman-Diaconis rule is used to regulate size of blocks depending on sample size (100,500,1000)

```{r, warning=FALSE}
set.seed(123)  
sample_sizes <- c(100,500,1000) 
for (sample_size in sample_sizes)
  {
x <- runif(sample_size)
simulated_data <- rGI0(x, p_alpha = -1, p_gamma = 1, p_Looks = 1)

# Bin Width with Freedman-Diaconis rule
iqr <- IQR(simulated_data)
bin_width <- 2 * iqr / (sample_size^(1/3))
bin_w<- sprintf("%.1f", bin_width)
num_bins <- ceiling((max(simulated_data) - min(simulated_data)) / bin_width)
num_bins<- nclass.FD(simulated_data)
x_range = c(0,20)
hist(simulated_data,  breaks = num_bins,
    main = paste("Simulated Data, sample size - ",sample_size,", num beans - ",num_bins) ,
    xlab = "x", ylab = "Frequency", probability = TRUE, col = "lightblue", xlim=c(0,20))
    lines(density(simulated_data), col = "darkblue",lwd=5)
}

```
<h1>3. Cullen-Frey Plots</h1>

Cullen-Frey plots are used to compare GI0 distribution with "canonical" models. For the analysis "fitdistrplus" package is used.

Cullen-Frey plots helps understanding the distribution's characteristics and evaluating how they related to canonical statistical models. This method was applied to both the simulated data generated from the GI0 distribution, uniform distribution, and normal distribution.

In "simulated data" plot observation are scattered and bootstrapping values are shown as a blue spot in the middle. They all are in the area of beta distribution, however as orange spots are not concentrated, it means that data does not fit well the beta model. If compare it to the normal and uniform distributions - observations are concentrated around bootstrapping values and bootstrapping values covers the signs for uniform (triangle) and normal (star) distributions correspondingly.
It also could be seen from skewness and kurtosis values.


```{r, warning=FALSE}
#Cullen-Frey plots
library(fitdistrplus)

# Create Cullen-Frey plots with specified titles
descdist(simulated_data, boot = 100)
mtext("Simulated Data")


descdist(x, boot = 100)
mtext("Uniform Data")

descdist(rnorm(1000), boot = 100)
mtext("Normal Data")


```
